### 2023-04-08 main/chapter_experiment.tex
some on pre-accelerators; split run3 upgrade

% does not stand alone
The LHC does not stand alone.
It implements the final boosts and collisions of particles, but requires
earlier stages of CERN's systems to inject the protons at suitably high
energies for it to continue their acceleration.
In brief, the accelerator chain flows as follows:
% Accelerator complex
\begin{enumerate}
% TODO where the protons come from
\item The protons come from
\end{enumerate}
This sequence is a subset of CERN's full accelerator complex,
which is illustrated in
Figure~\ref{fig:experiment_cern_complex}~\cite{Haffner:1621894}.

>>>>

% does not stand alone
The LHC does not stand alone.
It implements the final boosts and collisions of particles, but requires
earlier stages of CERN's systems to inject the protons at suitably high
energies for it to continue their acceleration.
In brief, the accelerator chain flows as follows:
% Accelerator complex
\begin{enumerate}
% where the protons come from
\item The protons come from hydrogen gas, which is initially ionised into a
plasma of protons and $e^-$ ions in the strong electric fields of a
duoplasmotron ion source, in which the protons are initially accelerated by
electric fields towards a cathode~\cite{lhc50years2013}.
% linac2
\item After ionisation, a linear accelerator named Linac2 provides the first
major proton acceleration in the chain, reaching $50\,\eV[M]$
energies~\cite{lhc50years2013}.
% PSB
\item Secondly, the Proton Synchrotron Booster (PSB) accelerates these protons
to $1.4\,\eV[G]$ energies.
The PSB comprises four rings with $160\,\mathrm{m}$
circumference~\cite{linac4protonsource,lhc50years2013}.
% PS
\item Thirdly, protons from the PSB are accelerated by the
Proton Synchrotron (PS) to $25\,\eV[G]$ energies~\cite{lhc2000psInjector}.
TODO
% SPS
\item SPS
% LHC
\item LHC
\end{enumerate}
This sequence is a subset of CERN's full accelerator complex,
which is illustrated in
Figure~\ref{fig:experiment_cern_complex}~\cite{Haffner:1621894}.
Each of these accelerators was upgraded from its previous operations to achieve
the performance demanded by designs for the LHC's high energy and luminosity.

Note that this sequence describes the LHC accelerator complex as it was for
operations that generated the data used in this thesis, in `Run~2' of the LHC.
It has since been upgraded.
For example, since the year 2020 (for Run~3 onwards), Linac2 is replaced by a
new linear accelerator named Linac4,
which accelerates $\mathrm{H}^-$ ions to $160\,\eV[M]$;
these $\mathrm{H}^-$ ions are later stripped of their electrons,
to bare protons by passage through a carbon foil.
Linac4 will supply higher luminosities for future operations of the accelerator
complex~\cite{linac4techreport,linac4protonsource}.
The PSB is since also upgraded to reach $2\,\eV[G]$ energies.
